### 2020-04-21 jss/jss-article-rnw/article.Rnw
Added template for mini sections in .Rnw form

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}
\usepackage{import}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Shannon K. Gallagher\\ National Institute of\\ Allergy and Infectious Diseases
   \And Benjamin P. Leroy\\Carnegie Mellon University}
\Plainauthor{Shannon K. Gallagher, Benjamin P. Leroy}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{\proglang{EpiCompare}: Working title}
\Plaintitle{EpiCompare}
\Shorttitle{\prolang{EpiCompare}}

%% - \Abstract{} almost as usual
\Abstract{
  This short article illustrates how to write a manuscript for the
  \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
  Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
  it is recommended to keep the {\LaTeX} code as simple as possible,
  i.e., avoid inclusion of packages/commands that are not necessary.
  For outlining the typical structure of a JSS article some brief text snippets
  are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
  discussing count data regression in \proglang{R}. Editorial comments and
  instructions are marked by vertical bars.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}

\SweaveInput{chapters/ch1-intro.Rnw}

\section{Purpose and Details}
\SweaveInput{chapters/ch2-vis.Rnw}
\SweaveInput{chapters/ch3-compare.Rnw}
\SweaveInput{chapters/ch4-convert.Rnw}

\section{In practice}
\SweaveInput{chapters/ch5-ex1.Rnw}
\SweaveInput{chapters/ch6-ex2.Rnw}

\SweaveInput{chapters/ch7-discussion.Rnw}

\section{references}

\bibliography{EpiCompare}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% -----------------------------------------------------------------------------

\end{document}