### 2020-04-21 jss/jss-article-rnw/chapters/ch2-vis.Rnw

% !Rnw root = ../article.Rnw

\subsection[Visualize]{Visualize} \label{sec:purpose-vis}