### 2020-04-21 jss/jss-article-rnw/chapters/ch5-ex1.Rnw

% !Rnw root = ../article.Rnw

\subsection[Case Study 1]{Case Study 1} \label{sec:ex1}

EXAMPLE 1