### 2020-04-22 jss/jss-article-rnw/chapters/app1-proofs.Rnw

% !Rnw root = ../article.Rnw

\section[Proof of Theorem]{Proof of Theorem} \label{app:proof}

\begin{proof}
\cite{harko2014} provide an analytical solution for the Kermack and McKendrick equations (Eq. \eqref{eq:sir-ode}) by reparameterizing the ODEs so that $\mathcal{S}(u) = S(t)$, $\mathcal{I}(u) = S(t)$, and $\mathcal{R}(u) = R(t)$ for $0< u_T < 1$ with
\begin{align}\label{eq:harko-odes}
\mathcal{S}(u) &= S(0)u\\
\mathcal{I}(u) &= N - R(0) + NR_0^{-1}\log u - S(0)u \nonumber\\
\mathcal{R}(u) &= R(0) - NR_0^{-1} \log u, \nonumber
\end{align}
and $u$ and t are related by the following integral,
\begin{align*}
    t &= \int_{u}^1 \frac{N}{\beta \tau (N - R(0) + R_{0}^{-1} \log \tau - S(0)\tau)}d\tau \\
    &= \int_{u}^1 \frac{1}{\beta f(S(0), R(0), N, R_0, \tau)} d \tau\\
    &= \int_{u}^1 \frac{1}{\beta f(\tau)} d\tau,
\end{align*}
where we have made the denominator of the integral a function of $N$, the initial values, $R_0$, and $\tau$, which we further condense to $f(\tau)$ for brevity.
Then for a given $t$ we want to find $s$ such that $(S_1(t), I_1(t), R_1(t)) = (S_2(s), I_2(s), R_2(s))$.  Or equivalently, for a fixed $u$ want to find $v$ such that  $\mathcal{S}_1(u) = \mathcal{S}_2(v)$ and then the corresponding $t$ and $s$ are given by
\begin{align*}
    t & = \int_{u}^1 \frac{1}{\beta_1 f(\tau)} d\tau \\
    s & = \int_{v}^1 \frac{1}{\beta_2 f(\tau)} d\tau.
\end{align*}
Note that since the equations in Eq. \eqref{eq:harko-odes} are functions of the initial values and $R_0$, then $u = v$. We then can find a relation for $s$,
    \begin{align*}
    s & = \int_{u}^1 \frac{1}{\beta_2 f(\tau)} d\tau  \\
    & = \int_{u}^1 \frac{1}{a\beta_1 f(\tau)} d\tau \\ 
    &= \frac{1}{a}\int_{u}^1 \frac{1}{\beta_1 f(\tau)} d\tau \\
    &= \frac{1}{a}t.
\end{align*}
\end{proof}