### 2020-04-30 jss/jss-article-rnw/chapters/ch4-orientation.Rnw

% !Rnw root = ../article.Rnw

\ben{This potentially could similar to that done in EpiModel's \citep{Jenness2018} orientation section, but I'm not sure - it might also depend upon which journal we aim to publish in. I also expect that this section will mirror how we present the pipeline in the Literature review.}