### 2020-04-30 jss/jss-article-rnw/chapters/ch4-orientation.Rnw
added comments relative to the outline of the paper

% !Rnw root = ../article.Rnw

\ben{This potentially could similar to that done in EpiModel's \citep{Jenness2018} orientation section, but I'm not sure - it might also depend upon which journal we aim to publish in. I also expect that this section will mirror how we present the pipeline in the Literature review.}

---

\ben{We also called this section ``Presenting EpiCompare'', not sure if we want to do that verse ``Package Orrientation'', but we'll figure it out.}

---

\ben{Also, note that the original sections proposed for this section and the Literature Review were:
\begin{enumerate}
\item Data collection
\item Exploratory analysis
\item ``Modelling'' $\rightarrow$ simulations
\item Assessment and comparisons
\end{enumerate}
}