### 2020-05-14 jss/jss-article-rnw/chapters/ch3.1-r0.Rnw
clean up of later section

% !Rnw root = ../article.Rnw

\subsection{Motivation through $R_0$}

$R_0$, the initial reproduction number, has been called the ``most important quantity in epdiemiology'' \citep[\ben{are you ok being cited here?}][]{Gallagher2020}. To see the importance of $R_0$, one need only read the newspaper \citep{Fisher} or look at the length of the table of estimated $R_0$ quantities for COVID-19 \citep[\ben{at shannon: is this an ok reference or do you want a better one?}][]{Aronson2020}. $R_O$ is also, maybe, the most famous \textit{time-free} numerical summary of an epidemic, and is most commonly associated with the Susceptible-Infectious-Recovered (SIR) data/models. The SIR framework, composed of data and statistical models where individuals pass from Susceptible to Infectious to Recovered states is a common, if not the most common, modeling framework in infectious disease epidemiology (seen in examples of recent published works in \cite{FILL IN}).

In regards to traditional $X$ vs. time plots, $R_0$ is difficult to visualize. For example, in a SIR simulation with two models have the same value of $R_0$, one can see in Fig. \ref{fig:different-scales-standard} that there is no way of knowing that from looking at the graph.

<<echo = F>>=
# \@Shannon: Make this figure reproducible here plz
# Also --- my vote is that we do this with points and lines? ~ Ben
@

\begin{figure}[!ht]
%% Code to reproduce is in inst/paper_figs/different-time-scales.R
    \centering
    \includegraphics[width = .5\textwidth]{images/diff-time-standard.pdf}%\includegraphics[width = .5\textwidth]{images/diff-time-ternary.pdf}
    \caption{Bivariate view of \# in each state vs. time. Both data sets are generated using the same value of $R_0 = 2.8$ but have different values of $\beta$ and $\gamma$.\ben{We probably need to define $\beta/\gamma$.}}
    \label{fig:different-scales-standard}
\end{figure}