### 2020-11-05 jss/jss-article-rnw/chapters/outline.Rnw

% !Rnw root = ../article.Rnw

\begin{enumerate}

\item Intro

\begin{enumerate}

\item Lots of epidemic \textit{infectious disease} models.  How do we handle them all

\item Pipeline of analysis

\item past work - many things under the wider umbrella of compartment models / multi state models (where one state is susceptible and one state is infectious)

\begin{itemize}
\item surveillance  -- one of the first major infectious disease R packages ED (kinda like it's own whole pipeline?)

\item pomp, epinet -- standardized frameworks of large classes of infectious disease models - modeling and simulation

\item MIDAS has done some work to catalogue different tools and models post- processing and comparison

\item Bring in tidyverse - standardization (plz throw spew a bone) pre-process, EDA and post processing

\item Patient vs. aggregate data (data pre-processing and EDA)

\item conformal bands and visualization for model assessment (skree plots, your hurricanes)

\end{itemize}
\end{enumerate}

\item EpiCompare qualities

\begin{enumerate}

\item standardizing outputs

\item TERNARY PLOTS

\item simple ode-style (also tidy-style) abms

\item time variant and analysis and conformal prediction bands

\end{enumerate}

\item Tour of EpiCompare

\begin{enumerate}
\item Highlights of Hagelloch
\item EDA by class
\item Simulation
\item Comparison with EpiNet model
\end{enumerate}

\end{enumerate}