### 2020-11-20 jss/jss-article-rnw/article.Rnw
Finally getting words on paper

\documentclass[article,shortnames]{jss} 
%^note: Ben changed this to get shortnames for references ("shortnames")

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}
\usepackage{import}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}
%\newcommand{\sup}{\mathop{\mathrm{sup}}}
%\newcommand{\inf}{\mathop{\mathrm{inf}}}

\usepackage{amsfonts,amsmath,amssymb,amsthm}
%\newtheorem{theorem}{Theorem}

% note making commands
\newcommand{\ben}{\textcolor{blue}}
\newcommand{\skg}{\textcolor{red}}

% useful packages
\usepackage[normalem]{ulem} % for \sout

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Shannon K. Gallagher\\ National Institute of\\ Allergy and Infectious Diseases
   \And Benjamin LeRoy\\Statistics and Data Science Department\\Carnegie Mellon University}
\Plainauthor{Shannon K. Gallagher, Benjamin LeRoy}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
% EpiCompare: A tidy R package for comparing and assessing epidemics and epidimology models
\title{\proglang{EpiCompare}: Working title}
\Plaintitle{EpiCompare}
\Shorttitle{\prolang{EpiCompare}}

%% - \Abstract{} almost as usual
\Abstract{
We do something, possibly useful
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Shanon K. Gallagher\\
  Biostatistics Research Branch\\
  Division of Clinical Research\\
  National Institute of Allergy and Infectious Diseases\\
  5601 Fishers Lane\\
  Rockville, MD 20852\\
  E-mail: \email{shannon.gallagher@nih.gov}
}

\begin{document}
\SweaveOpts{concordance=FALSE}

\section{Outline}

\SweaveInput{chapters/outline.Rnw}

\SweaveInput{chapters/ch1-intro-v2.Rnw}

%\SweaveInput{chapters/ch1-intro.Rnw}

\section{Literature review} \label{sec:lit}
\SweaveInput{chapters/ch2-lit-review.Rnw}

\section{Reasons and tools for time-invariant analysis} \label{sec:time-invariant}
\SweaveInput{chapters/ch3-time-free.Rnw}

%\SweaveInput{chapters/ch2-vis.Rnw} % this file still exists
%\SweaveInput{chapters/ch3-compare.Rnw}
%\SweaveInput{chapters/ch4-convert.Rnw}

\section{Code Orientation}
\SweaveInput{chapters/ch4-orientation.Rnw}

\section{A tour of EpiCompare}
\SweaveInput{chapters/hagelloch-ex.Rnw}

\SweaveInput{chapters/ch7-discussion.Rnw}

\bibliography{EpiCompare}

\appendix

\SweaveInput{chapters/app1-proofs.Rnw}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% -----------------------------------------------------------------------------

\end{document}