### 2020-12-20 README.Rmd
attempt 1 for github action (dangerous) - check, covr and pkgdown all in 1

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300
)
```

# EpiCompare <a href = "https://skgallagher.github.io/EpiCompare/"><img src="man/figures/logo.png" align="right" alt="EpiCompare" width="120" /></a>

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/skgallagher/EpiCompare.svg?branch=master)](https://travis-ci.org/skgallagher/EpiCompare)
[![codecov](https://codecov.io/gh/skgallagher/EpiCompare/branch/master/graph/badge.svg)](https://codecov.io/gh/skgallagher/EpiCompare)
[![R-CMD-check](https://github.com/skgallagher/EpiCompare/workflows/R-CMD-check/badge.svg)](https://github.com/skgallagher/EpiCompare/actions)
<!-- badges: end -->

The goal of `EpiCompare` is to provide the epidemiology community with easy-to-use tools to encourage comparing and assessing epidemics and epidemiology models in a time-free manner. All tools attempt to adhere to `tidyverse`/`ggplot2` style to enhance easy of use.

Time free analysis allows for stronger comparison of epidemics and model based simulations avoiding different scaling and shifts of time that mess up time-based comparisons.

To achieve this goal, the package contains:

  - **Visualization tools** to visualize SIR epidemics and simulations from SIR models in a time-free manner using `ggtern`’s ternary plots and prediction bands. For agent-based SIR models we also provide visualization tools to let the user easily explore how different characteristics of the agents relate to different experiences in the epidemic.
  - General **comparison tools** to compare epidemics and epidemic models that have higher numbers of states (again in a time-free manner), allowing for the user to examine the differences between models through simulations, and if an epidemic is similar to a model through simulations and prediction bands.
  - **Conversion tools** to: 
    - Convert and then compare models from *standard epidemic packages* like `EpiModels`, `pomp`, as well as internal agent-based models, and epidemics in a common framework. 
    - Convert *agent-based information into aggregate* to compare in the aggregate framework described above.

## Installation

You can install the developmental version of EpiCompare from github using:

``` r
# install.packages("devtools")
devtools::install_github("skgallagher/EpiCompare")
```

```{r echo = F, warning = F, message = F}
devtools::load_all()
```
## Data

Description of data including in this package can be found in the data section of the [reference](https://skgallagher.github.io/EpiCompare/reference/index.html#section-data) page of the documentation website.

```{r eval = F, echo = F}
   Currently there are 13 data sets in which to explore the different `EpiCompare` functions.  These correspond to different sources of raw data along, agent vs. aggregate format, and emphases of different processes such as vital dynamics (e.g. birth and).

   Details of each can be found with `?{data}`.

  - `agent_sims` this is 50 x 3 x 188 array where entry (i,j,k) looks at the ith simulation, the jth statistic, and the kth agent.  The statistics are (inititial state (0/1/2), SMax, IMax).

  - `agent_sims_tidy` ("Simulations of Measles outbreaks for Hagelloch, Germany, 1861 (Tidy agent format)"): tidy version of `agent_sims` (9400 x 5), each row corresponds to an individual for a single simulation, and contains additional information on the individual's initial state, and SMax, IMax.

  - `EpiModel_agg_bd` Example output from the `EpiModel package` for an ICM with birth and death rates.  The population $N$ is non-constant.

  - `EpiModel_det` output from the `EpiModel` package for a deterministic model (aggregate)

  - `EpiModel_icm` Output from the `EpiModel` package for a stochastic ICM (aggregate)

  - `hagelloch_agents` -- One row is a "sufficient" statistic for each agent's infection.  Each agent's infection is uniquely identified by an initial state, max time before infection (or T), and max time before recovery (or T).  For the states, 0 = S, 1 = I, 2 = R.

  - `hagelloch_aug_births` Like Hagelloch raw but augmented with births (it already had deaths).  Five fake people have been added who join the population over time

  - `hagelloch_raw` -- One row is an agent.  This is imported from the `surveillance` R package and the variable descriptions are found [here](https://rdrr.io/rforge/surveillance/man/hagelloch.html) where it is originally labeled `hagelloch.df`.  We have renamed it here to help distinguish it from the other data sets we derive from it.

  - `hagelloch_raw2` Like Hagelloch raw but with additional, fake people with NA values for `tI` and `tR`

  - `hagelloch_sir`  -- One row is a state of $(t, s_t, i_t, r_t)$ where $s_t + i_t + r_t = N$ for $t = 0, \dots, T=94$

  - `pomp_arr` Example SIR simulation output of class `array` from the `pomp` packagee

  - `pomp_df` Example SIR simulation of class `data.frame` from the `pomp` package

  - `pomp_pomp` Example SIR simulation output of class `pomp` from the `pomp` package

  - `pomp_sir` Example output from the `pomp` package.
```
## Example

```{r message = F, warning = F}
library(tidyr)
library(ggplot2)
library(dplyr)
library(ggtern)
library(EpiCompare)
```
The following example comes from a Measles outbreak in Hagelloch, Germany in 1861. We have data on each child (agent) in the town.
```{r}
hagelloch_raw %>%
  ggplot(aes(y = tI, z = tR)) +
    geom_aggregate() + 
    coord_tern() +
    labs(x = "S", y = "I", z = "R", title = "Town Analysis") +
    theme_sir()
```

Previous work has suggested that the class (`CL`) the student was in effected how the experienced the outbreak. The below figure shows differences in the outbreak relative to this grouping.
```{r}
hagelloch_raw %>% 
  rename(`school grade` = CL) %>%
  group_by(`school grade`) %>%
  summarize(`number of students` = n())

hagelloch_raw %>%
   ggplot(aes(y = tI, z = tR, color = CL)) +
       geom_aggregate() + 
       coord_tern() +
       labs(x = "S", y = "I", z = "R",
            color = "School Grade",
            title = "Town analysis by grade") +
       theme_sir()
```

### Simulate SIR data

```{r}
n_sims <- 100
n_time_steps <- 100
beta <- .1
gamma <- .03
init_SIR <- c(950, 50, 0)

out <- simulate_SIR_agents(n_sims = n_sims,
                           n_time_steps = n_time_steps,
                           beta = beta, gamma = gamma,
                           init_SIR = init_SIR)

df_groups <- out %>% dplyr::group_by(sim) %>%
    agents_to_aggregate(states = c(tI, tR)) %>%
    rename(S = X0, I = X1, R = X2)

df_groups %>% ggplot() +
    geom_prediction_band(aes(x = S, y = I, z = R, sim_group = as.numeric(sim)),
                         alpha = .2, fill = "blue", color = "blue") +
    geom_line(aes(x = S, y = I, z = R, group = sim), alpha = .1) +
    coord_tern() +
    theme_sir()
```

## Package Creation Notes:
1. For writing code that works with `tidyverse` 1.0 vs `tidyverse` <= 0.8.3. We followed ideas found in [tidyr: in-packages](https://tidyr.tidyverse.org/articles/in-packages.html), for the code, and - when working with Travis CI (using a matrix for multiple builds) - we leverage ideas in [tidyverse travis on github](https://github.com/tidyverse/design/issues/95) and [tidyverse principles](https://principles.tidyverse.org/changes-multivers.html).
2. For writing your own `geom`s and `stat`s that works with `ggtern` (which are generally restricted), the following 2 stack-exchange articles helped use do so with ease:
    a. [stack-exchange: personal code idea](https://stackoverflow.com/questions/57175114/ternary-plot-scaling-opacity-across-groups)
    b. [stack-exchange: being able to access ggtern's element write away](https://stackoverflow.com/questions/31589479/how-to-fix-no-item-called-packagepkg-on-the-search-list-without-using-libra#comment51172891_31589479)

    c. Finally, we've also leveraged ideas from [R-devel: on avoiding problems with `:::`](https://stat.ethz.ch/pipermail/r-devel/2013-August/067210.html) in `R/aaa.R` to overcome messages from CRAN relative to this hack (using `:::`). For some reason - when documenting for `pkgdown` website, we need to do `library(ggtern); EpiCompare:::update_approved_layers()`
3. `geom_prediction_band` required not just `compute_group` but `compute_layer` - there is very little documentation on how to approach this correctly. Basically - there are problems when the `compute_group` wants to make multiple `pieces`/`groups` - and it is similar to the problem that if you do something like `aes(color = var1, group = var2)` you may actually want to do `aes(color = var1, group = paste(var1, var2))`, if there are the same `var2` values across different `var1` values but they don't mean they should be grouped together.
4. Now that `R` have come out with version >= 4.0.0, we now need to call `.S3method("method", "class")` to define the connection for `S3` methods (e.g. `method.class` function), which we have for the `contained` function. 
5. **Does 2 wrongs make a right?** As of 9/23 `ggtern` had an [issue](https://bitbucket.org/nicholasehamilton/ggtern/issues/13/ggtern-breaks-ggplots-plot) that it messed with `ggplot2`'s legends when loaded (it over-wrote the `print.ggplot` and other functions). To "correct" this, we've over-wrote `ggtern`'s `print.ggplot` to correct this problem when not producing ternary plots (code in `aaa.R`). 
6. Useful Rstudio shortcuts for `Roxygen2`: (a) create `Roxygen2` comments template with `option` + `command` + `shift` + `R` (b) In `Roxygen2` comments do `control` + `shift` + `/` to format relative to 80 char limit.
7. [stack overflow](https://stackoverflow.com/questions/11285496/r-cmd-check-warning-functions-methods-with-usage-in-documentation-object-bu) post on how to pass `check` for `rownames<-.tidy_dist_mat`.

## Contributors

- Shannon Gallagher ([`skgallagher`](https://github.com/skgallagher))
- Benjamin LeRoy ([`benjaminleroy`](https://github.com/benjaminleroy))