### 2020-12-24 README.Rmd
more detailed description of geom_prediction_band inputs values

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300
)
```

# EpiCompare <a href = "https://skgallagher.github.io/EpiCompare/"><img src="man/figures/logo.png" align="right" alt="EpiCompare" width="120" /></a>

<!-- badges: start -->
[![R-CMD-check, code coverage](https://github.com/skgallagher/EpiCompare/workflows/R-CMD-check%20and%20code%20coverage/badge.svg)](https://github.com/skgallagher/EpiCompare/actions)
<!-- [![Travis build status](https://travis-ci.org/skgallagher/EpiCompare.svg?branch=master)](https://travis-ci.org/skgallagher/EpiCompare) -->
[![codecov](https://codecov.io/gh/skgallagher/EpiCompare/branch/master/graph/badge.svg)](https://codecov.io/gh/skgallagher/EpiCompare)

<!-- badges: end -->

The goal of `EpiCompare` is to provide the epidemiology community with easy-to-use tools to encourage comparing and assessing epidemics and epidemiology models in a time-free manner. All tools attempt to adhere to `tidyverse`/`ggplot2` style to enhance ease of use.

Additionally, `EpiCompare` provides tools for time invariant analysis.   This allows for 'fairer' comparison of epidemics and model based simulations by avoiding different scaling and shifting of time that can confound time-based comparisons.

To achieve these goals, the package contains:

  - **Visualization tools** to visualize SIR epidemics and simulations from SIR models in a time-free manner using `ggtern`’s ternary plots and prediction bands. For agent-based SIR models we also provide visualization tools to let the user easily explore how different characteristics of the agents relate to different experiences in the epidemic.
  - General **comparison tools** to compare epidemics and epidemic models that have higher numbers of states (again in a time-free manner), allowing for the user to examine the differences between models through simulations, and if an epidemic is similar to a model through simulations and prediction bands.
  - **Conversion tools** to: 
    - Convert and then compare models from *standard epidemic packages* like `EpiModels`, `pomp`, as well as internal agent-based models, and epidemics in a common framework. 
    - Convert *agent-based* information into *aggregate* to compare in the aggregate framework described above.

## Installation

You can install the developmental version of EpiCompare from github using:

``` r
# install.packages("devtools")
devtools::install_github("skgallagher/EpiCompare")
```

```{r echo = F, warning = F, message = F}
devtools::load_all()
```
## Data

Description of data including in this package can be found in the data section of the [reference](https://skgallagher.github.io/EpiCompare/reference/index.html#section-data) page of the documentation website.

## Example

```{r message = F, warning = F}
library(tidyr)
library(ggplot2)
library(dplyr)
library(ggtern)
library(EpiCompare)
```
The following example comes from a Measles outbreak in Hagelloch, Germany in 1861. We have data on each child (agent) in the town.
```{r}
hagelloch_raw %>%
  ggplot(aes(y = tI, z = tR)) +
    geom_aggregate() + 
    coord_tern() +
    labs(x = "S", y = "I", z = "R", title = "Town Analysis") +
    theme_sir()
```

Previous work has suggested that the class (`CL`) the student was in effected how the experienced the outbreak. The below figure shows differences in the outbreak relative to this grouping.
```{r}
hagelloch_raw %>% 
  rename(`school grade` = CL) %>%
  group_by(`school grade`) %>%
  summarize(`number of students` = n())

hagelloch_raw %>%
   ggplot(aes(y = tI, z = tR, color = CL)) +
       geom_aggregate() + 
       coord_tern() +
       labs(x = "S", y = "I", z = "R",
            color = "School Grade",
            title = "Town analysis by grade") +
       theme_sir()
```

### Simulate SIR data

```{r}
n_sims <- 100
n_time_steps <- 100
beta <- .1
gamma <- .03
init_SIR <- c(950, 50, 0)

out <- simulate_SIR_agents(n_sims = n_sims,
                           n_time_steps = n_time_steps,
                           beta = beta, gamma = gamma,
                           init_SIR = init_SIR)

df_groups <- out %>% dplyr::group_by(sim) %>%
    agents_to_aggregate(states = c(tI, tR)) %>%
    rename(S = X0, I = X1, R = X2)

df_groups %>% ggplot() +
    geom_prediction_band(aes(x = S, y = I, z = R, sim_group = as.numeric(sim)),
                         alpha = .2, fill = "blue", color = "blue") +
    geom_line(aes(x = S, y = I, z = R, group = sim), alpha = .1) +
    coord_tern() +
    theme_sir()
```

## Package Creation Notes:

**We're transferring to ~github actions~ and away from Travis CI. Thanks Travis CI for the long run (During undergrad - probably around 2015, I got introduced to Travis CI and it has been a really great tool and CIs in general are great tools). Sadly, open source packages (like ours) no longer gets infinite free resources on Travis. [Dean Attali](https://deanattali.com/blog/migrating-travis-to-github/) and [ROpenSci](https://ropensci.org/technotes/2020/11/19/moving-away-travis/) have blog posts on the situation. As such, if you're looking to learn from our mistakes from Travis, then the comments below stop making sense after December 15th, 2020.

1. For writing code that works with `tidyverse` 1.0 vs `tidyverse` <= 0.8.3. We followed ideas found in [tidyr: in-packages](https://tidyr.tidyverse.org/articles/in-packages.html), for the code, and - when working with Travis CI (using a matrix for multiple builds) - we leverage ideas in [tidyverse travis on github](https://github.com/tidyverse/design/issues/95) and [tidyverse principles](https://principles.tidyverse.org/changes-multivers.html). **This is no longer done (was removed 22 December 2020), as the rest of `tidyrverse` has moved on and now requires `tidyr >1.0.0`.
2. For writing your own `geom`s and `stat`s that works with `ggtern` (which are generally restricted), the following 2 stack-exchange articles helped use do so with ease:
    a. [stack-exchange: personal code idea](https://stackoverflow.com/questions/57175114/ternary-plot-scaling-opacity-across-groups)
    b. [stack-exchange: being able to access ggtern's element right away](https://stackoverflow.com/questions/31589479/how-to-fix-no-item-called-packagepkg-on-the-search-list-without-using-libra#comment51172891_31589479)

    c. Finally, we've also leveraged ideas from [R-devel: on avoiding problems with `:::`](https://stat.ethz.ch/pipermail/r-devel/2013-August/067210.html) in `R/aaa.R` to overcome messages from CRAN relative to this hack (using `:::`). For some reason - when documenting for `pkgdown` website, we need to do `library(ggtern); EpiCompare:::update_approved_layers()`
3. `geom_prediction_band` required not just `compute_group` but `compute_layer` - there is very little documentation on how to approach this correctly. Basically - there are problems when the `compute_group` wants to make multiple `pieces`/`groups` - and it is similar to the problem that if you do something like `aes(color = var1, group = var2)` you may actually want to do `aes(color = var1, group = paste(var1, var2))`, if there are the same `var2` values across different `var1` value, but should not necessarily be grouped together.
4. Now that `R` has come out with version >= 4.0.0, we now need to call `.S3method("method", "class")` to define the connection for `S3` methods (e.g. `method.class` function), which we have for the `contained` function. 
5. **Do two wrongs make a right?** As of 9/23 `ggtern` had an [issue](https://bitbucket.org/nicholasehamilton/ggtern/issues/13/ggtern-breaks-ggplots-plot) that it messed with `ggplot2`'s legends when loaded (it over-wrote the `print.ggplot` and other functions). We've over-written `ggtern`'s `print.ggplot` to correct this problem when not producing ternary plots (code in `aaa.R`). 
6. Useful Rstudio shortcuts for `Roxygen2`: (a) create `Roxygen2` comments template with `option` + `command` + `shift` + `R` (b) In `Roxygen2` comments do `control` + `shift` + `/` to format relative to 80 char limit.
7. [stack overflow](https://stackoverflow.com/questions/11285496/r-cmd-check-warning-functions-methods-with-usage-in-documentation-object-bu) post on how to pass `check` for `rownames<-.tidy_dist_mat`.
8. Transferring from Travis CI to github actions. We only use a single workflow file (although we use code ideas found in: [check-standard/`usethis::use_github_action_check_standard()`](https://github.com/r-lib/actions/blob/master/examples/check-standard.yaml)), [test-coverage/`usethis::use_github_action("test-coverage")`](https://github.com/r-lib/actions/blob/master/examples/test-coverage.yaml). In our `.github/workflows/R-CMD-check-coverage.yaml` you'll find **a** (potentially not optional) approach to preform our complex checking approach (which tries to copy the ideas in our [old travis file](https://github.com/skgallagher/EpiCompare/blob/e1298f799d9248bb330885ba5e7b0fa17ea2c83e/.travis.yml)). The github action easily creates a *larger* matrix that allows us to run on all standard OS, R versions and our "tidyr-current" vs "tidyr-old" split. Our approach with 1 github action `yaml` file makes us only look at 8 matrix options - it's possible that, in the future, we'll go back to 3 files. We currently compile our `pkgdown` site on our own computers and then push - and will continue to do so due to complications with `reticulate` and `python` packages on the virtual machines in github actions.

## Contributors

- Shannon Gallagher ([`skgallagher`](https://github.com/skgallagher))
- Benjamin LeRoy ([`benjaminleroy`](https://github.com/benjaminleroy))