### 2021-04-15 jss/EpiCompare-paper/EpiCompare-paper.tex
ben responses to comments on section 3 - minus last section which we will discuss in person (first round of edits after completition of section 3)

\documentclass[
  shortnames]{jss}

\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
Shannon K. Gallagher\\Biostatistics Research Branch\\
National Institute of Allergy\\
and Infectious Diseases \And Benjamin LeRoy\\Dept. of Statistics \& Data Science\\
Carnegie Mellon University
}
\title{Time invariant analysis of epidemics with \pkg{EpiCompare}}

\Plainauthor{Shannon K. Gallagher, Benjamin LeRoy}
\Plaintitle{Time invariant analysis of epidemics with EpiCompare}
\Shorttitle{\pkg{EpiCompare}}

\Abstract{
We present \pkg{EpiCompare}, an \proglang{R} package that suppliments
and enhances current infectious disease analysis pipelines and
encourages comparisons across models and epidemics. A major contribution
of this work is the set of novel \textit{time-invariate} tools for model
and epidemic comparisons - including time-invariate prediction bands.
\pkg{EpiCompare} embraces \proglang{R}'s \textit{tidy} coding style to
make adoption of the package easier and analysis faster. This paper
provides an overview of both the tools in and intuition behind
\pkg{EpiCompare} and a thorough demonstrating of the tools through a
detailed example of a full data analysis pipeline.
}

\Keywords{keywords, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, not capitalized, Java}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    Shannon K. Gallagher\\
    Biostatistics Research Branch\\
  National Institute of Allergy\\
  and Infectious Diseases\\
    5603 Fishers Lane\\
Rockville, MD 20852\\
  E-mail: \email{shannon.gallagher@nih.gov}\\
  URL: \url{http://skgallagher.github.io}\\~\\
      Benjamin LeRoy\\
    Dept. of Statistics \& Data Science\\
  Carnegie Mellon University\\
    5000 Forbes Ave.\\
Pittsburgh, PA 15213\\
  E-mail: \email{bpleroy@andrew.cmu.edu}\\
  URL: \url{https://benjaminleroy.github.io/}\\~\\
  }

% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{xcolor}
\usepackage{rotating}

\usepackage{amsmath} \usepackage{amssymb} \usepackage{amsthm} \usepackage{afterpage} \usepackage[normalem]{ulem}

\begin{document}

\newcommand{\shannon}[1]{\textcolor{orange}{#1}}
\newcommand{\ben}[1]{\textcolor{violet}{#1}}

\newtheorem{theorem}{Theorem}

\section[Intro]{Introduction}\label{sec:intro}

The recent (and on-going) COVID-19 global pandemic has galvanized public
interest in understanding more about infectious disease modeling and has
highlighted the usefulness of research in the area of infectious disease
epidemiology. Infectious diseases inflict enormous burdens on the world:
millions of lives lost and trillions of dollars spent yearly. Infectious
disease models typically attempt to do one or more of the following: 1)
predict the spread of current and future epidemics
\citep[e.g. flu prediction][]{Biggerstaff2016}, 2) analyze past and
current epidemics to increase scientific knowledge
\citep[e.g. historical measle outbreaks][]{Neal2004}, and 3) forecast or
project epidemic scenarios under pre-specified parameters
\citep[e.g.][]{ferguson2020}. At the same time, descriptive statistics
and visualizations from universities, many branches and levels of
government, and news organizations are an important first step of the
process
\textcolor{violet}{as has been seen in the current COVID-19 epidemic}\citep{dong2020,cdc-covid-tracker2021,wp-covid-tracker2021}.
\footnote{\textcolor{violet}{[Ben says: probably should have a conclusion sentence here - seems to end abruptly. *This is less so the case now.]}}

With the many visualization and exploratory tools, models and modeling
paradigms, and reviews and comparisons in the literature and through the
MIDAS (Models of Infectious Disease Agent Study) network
\citep{midasnetwork2021}, this field has a lot of devices to aid an
individual practitioner decide the correct approach. For example,
\proglang{R} packages such as \pkg{surveillance}, \pkg{EpiModel}, and
\pkg{pomp} have all made significant steps in standardizing the flow of
the data analysis pipeline for epidemic modeling through digitizing data
sets, making accessible statistical models, and providing a plethora of
educational material for both coding novices and experts alike
\citep{surveillance2017,Jenness2018,King2016}.

At the same time, analysis packages often only address a specific
portion of the analysis
pipeline\textcolor{violet}{\sout{, for instance focusing on certain types of models.}}
\textcolor{violet}{These m}odeling
tools\textcolor{violet}{\sout{, which}} usually require learning
package-specific syntax\textcolor{violet}{\sout{,}} and often don't
provide easy ways to compare and assess their models on new data.
Moreover, exploring\textcolor{violet}{, \sout{and}} modeling
\textcolor{violet}{and comparing} epidemics require transforming and
\textit{tidying} data in different ways. To fill these gaps, we present
our \proglang{R} package \pkg{EpiCompare}. Our package's primary focus
is to aid and advance research in the area of comparison and assessment
of epidemic and epidemiological models. In Figure \ref{fig:pipeline}, we
illustrate the data analysis pipeline of infectious diseases as 1) data
pre-processing, 2) exploratory data analysis (EDA), 3) modeling and
simulating, 4) post-processing, and 5) comparison and assessment; where
each previous part of the pipeline influences the next. \pkg{EpiCompare}
provides tools to aids practitioners in all areas of this pipeline.

\begin{figure}[!ht]
    \centering
    \includegraphics[width = 1\textwidth]{images/pipeline1.png}
    \caption{An idealized epidemiological data analysis pipeline.}
    \label{fig:pipeline}
\end{figure}

\pkg{EpiCompare} also emphasizes the value of analyzing epidemics in a
\textit{time-invariant} way. Epidemics, despite by definition being a
process that evolves over time, often need to be compared in a way not
constrained to initial times or time scales to understand the processes
at play.
\textcolor{violet}{Time-invariant analysis can also make it easier to compare state-space models in a more global, holistic fashion. \sout{Moreover, m} M}any
current \textcolor{violet}{time-dependent} comparison tools for
state-space models (e.g.~SIR models)
\textcolor{violet}{\sout{highlight} examine} the proportion of
individuals in each state (at a given time) in a piece-wise / marginal
fashion. \textcolor{violet}{These \sout{This}}
approach\textcolor{violet}{es} may reduce the amount of connections that
can be seen, similar to projections of a multidimensional distribution
onto a single axis at a time. Tools in \pkg{EpiCompare} give the user
the ability to extend their toolkit to evaluate epidemics within a
time-invariant lens. The goal of \pkg{EpiCompare} is not to supplant
existing infectious disease modeling tools and software but, rather, is
a concerted effort to create standard and fair comparisons among models
developed for disease outbreaks and outbreak data.

This paper is broken up into the following sections; section
\ref{sec:time-invariant} motivates and showcases tools of time-invariant
analysis, section \ref{sec:overview} presents an outline of how
\pkg{EpiCompare} aids a practitioner in every step of the pipeline and
section \ref{sec:tour} provides a \textcolor{violet}{\sout{thorough}}
demonstrating of the tools through a detailed example of a full data
analysis pipeline.

\section[Time-invariant]{Motivation and tools for time-invariant
analysis}\label{sec:time-invariant}

\pkg{EpiCompare} delivers \textit{time-invariant} analysis by (1) taking
a global, not marginal view of how epidemics move through populations
and (2) by treating full epidemics as filamental trajectories. The
following section aims to highlight the strengths of
\textit{time-invariant} \textcolor{orange}{analysis} and define the
mathematical foundations that \pkg{EpiCompare}'s tools stand upon.

Mathematically, epidemics are complex objects. They can be hard to
assess and compare to one another due to the differences in the
diseases, the location where the outbreak occurs, how the affected
population reacts, and the time
\sout{related}\textcolor{orange}{related} features (including start of
the epidemic, speed of infection and more). Time-invariant analysis
makes different epidemics easier to compare by removing many time
dependent aspects of an epidemic. Instead,
\textcolor{orange}{time-invariant analysis} focuses
\textcolor{orange}{on the global pattern of an epidemic, via filamental trajectories,  and emphasizes the number of lives affected.}
\textcolor{violet}{[Ben wants to try this sentence again.]}

\subsection[motivating through R0]{Motivating time-invariant analysis
through the reproduction number \(R_0\)}\label{sec:r0}

Time-invariant analysis, as it appears in \pkg{EpiCompare},
\textcolor{orange}{bypasses} many difficulties \textcolor{orange}{in}
comparing different epidemics. With time-invariant analysis, comparing
the decades-long outbreak of HIV in the US to a 10 day outbreak of
norovirus on a cruise ship is \sout{still} possible. Time-dependent
problems can arise when estimating epidemiological parameters, including
the reproduction number \(R_0\).
\sout{\textcolor{violet}{We will use $R_0$ to motivate the usefulness of time-invariant analysis in this section.}}\footnote{\textcolor{orange}{I don't think this is a necessary sentence.}
  \textcolor{violet}{I still think it adds value to the story and I'm not sure people really read section titles that are long.}}

\(R_0\) is probably the most famous \sout{time-invariant} numerical
summary of an epidemic and is often associated with the
Susceptible-Infectious-Recovered (SIR) model \citep{hethcote2000}.
\(R_0\) is \sout{a one-number summary of a disease and is }defined as
the expected number of infections caused by a single infector who is
added to a completely susceptible population \citep{anderson1992}.
\textcolor{violet}{This \textcolor{orange}{definition has no mention of time and hence} means that $R_0$ is a time-invariant parameter\textcolor{orange}{. Yet $R_0$ is} estimated with time-\sout{based}\textcolor{orange}{dependent} data, which can make it a difficult quantity to estimate.}
\textcolor{orange}{For example, \cite{Gallagher2020} demonstrate how $R_0$ can be sensitive to time-\sout{based}\textcolor{orange}{dependent} parameters such as the beginning and end of an epidemic, two quantities that generally are}\textcolor{violet}{hard to define precisely.\sout{do not have precise definitions}}.
To demonstrate the difficulty of discerning \(R_0\) in
\textcolor{violet}{\sout{a}other}\footnote{\textcolor{violet}{I change this so we don't confused readers that we're going show the impact in tools beyond just the estimation itself.}}
time-dependent analysis, we first introduce \citet{Kermack1927}'s SIR
model. This model captures the transitions from one state to the next as
a system of ordinary differential equations, where \(N\) is the total
number of individuals, \(\beta\) is the rate of infection, and
\(\gamma\) is the rate of recovery,

\begin{align}\label{eq:sir-ode}
    S^\prime(t) &= -\frac{\beta S(t)I(t)}{N} \\
    I^\prime(t) &= \frac{\beta S(t)I(t)}{N} - \gamma I(t) \nonumber\\
    R^\prime(t) &= \gamma I(t) \nonumber.
\end{align}

From this
model,\textcolor{orange}{the reproduction number is the ratio of the infection rate to the recovery rate,}
\(R_0 = \beta/\gamma\)\sout{, aka the ratio of the infection rate compared to the recovery rate.}
\textcolor{violet}{From this definition, given 
\sout{Since}} \(\beta\) and \(\gamma\) are both rates,
\textcolor{violet}{it should be clear that} the ratio of the two,
\(R_0\), is a time-invariant
quantity.\footnote{\textcolor{violet}{I am trying to make it look like we are not repeating ourselves by saying $R_0$ is time-invariant.}}
\sout{ Once $R_0$ is estimated, practitioners can infer important epidemic quantities such as the total number of infections or the percent of a population needed to be vacccinated to stop the sustained spread of an epidemic.  Moreover, $R_0$ allows us to compare different diseases and different instances of outbreaks on the same scale. }\footnote{cool
  facts about r0, but not the central point}

\textcolor{violet}{[Ben says: It's unclear to me why we have a subtitle here - isn't it just more motivation of time-invariant anlysis with $R_0$? Also, I feel like the story is weak here. The point is to leverage $R_0$ to show the value of time-invariant analysis - this seems a bit more like just discussing properties of $R_0$. In the follow rewrite I use "[" and "]" to indicate that this is a section from your earlier draft.]}
\textcolor{orange}{[Shannon says: Tried to tie this better to the previous part since it's no longer a new section.  also highlighted tie to time-invariant analysis and $R_0$.  I also wanted to bring the punch line (overlapping epidemics = same r0) closer to the beginning so those who don't want to slog through mathematical details can get the takeaway.]}\textcolor{orange}{ Shannon tries again in \textcolor{blue}{blue}}

\sout{\textcolor{violet}{[Ben says: this paragraph needs to still be looked at. Also I'm not sure why this particular paragraph was changed so much at all. Currentlly, the way we present comparing these epidemic's $R_0$s isn't well grouped/motivated.]} \textcolor{orange}{Time-invariant analysis helps practitioners to more easily compare $R_0$ from different outbreaks.} For example, consider two epidemics generated from the Kermack and McKendrick SIR equations. The first epidemic has parameters $\beta_1, \gamma_1 = (0.8,0.4)$ and the second has $\beta_2, \gamma_2 = (0.64,0.32)$.  Both epidemics have populations of 1000 people with 10 individuals initially infected.  Additionally note that the two reproduction numbers are the same for each epidemic, $R_0 = 2=0.8/0.4 = 0.64/0.32$.  We plot the epidemics with traditional $state$ vs. $time$ plots}\footnote{This sentecne is out of place / doesn't connect with the other sentences.}.
\sout{In Fig. \ref{fig:different-scales-standard} we show the time-based paths for the $S$, $I$, and $R$ states for the first 15 days of observed data.  In this time-variant view, we may believe that \textcolor{violet}{epidemic} 1 has a larger $R_0$ than \textcolor{violet}{epidemic} 2 because the peak of infection occurs more quickly than in Epidemic 2.  On the other hand, we may believe \textcolor{violet}{epidemic} 2 has a larger $R_0$ because its unclear if the  number of infections in that \textcolor{violet}{epidemic} has not yet peaked at time 15.  In this time-variant view, we cannot determine if one epidemic has larger value of $R_0$}\footnote{\textcolor{violet}{This sentnce doesn't connect wit previous examples.}}.

\textcolor{blue}{Since $R_0$ is an important value, it would be helpful to have more intuitive ways of comparing one $R_0$ to another.  Usually numerical summaries of $R_0$ are presented, which while overall very helpful, may be confusing when presented along side epidemic data that are visualized in a traditional, time-dependent manner.}

\textcolor{blue}{For example, consider two epidemics generated from the Kermack and McKendrick SIR equations where both have the same value of $R_0$. The first epidemic has parameters $\beta_1, \gamma_1 = (0.8,0.4)$ and the second has $\beta_2, \gamma_2 = (0.64,0.32)$.  Both epidemics have populations of 1000 people with 10 individuals initially infected. An analysis may present an estimate of $\hat{R}_0 = 2$ alongside state vs. time plots like those shown in Figure \ref{fig:different-scales-standard}. The paths of the epidemics in the state vs. time view seem to differ from one another including having different infection peaks. From these traditional time-based plots, there is no intuitive way to conclude that these two epidemics have the same value of $R_0$.}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{Figs/unnamed-chunk-2-1}

}

\caption{\label{fig:different-scales-standard}Example of two epidemics with different $\beta$ and $\gamma$ paremeters but the same initial reproduction number $R_0$ = 2.  Both epidemics are generated from models with $N= 1000$ individuals with $S(0) = 990$ and $I(0) = 10$.}\label{fig:unnamed-chunk-2}
\end{figure}
\end{CodeChunk}

\textcolor{blue}{\pkg{EpiCompare} provides a time-invariant tool to visualize these epidemics in a more intuitive manner, at least in regards to comparing values of $R_0$.}
\sout{A time-invariant approach to visualizing epidemics, in comparison, allows us to directly compare $R_0$ from a single plot.}
\textcolor{violet}{For every time \sout{point} $t$ we have a \textcolor{orange}{point} $(S(t),I(t), R(t))$\sout{, so we can treat epidemics as a trajectory in this three-dimensional space, as we visual in the left subplot of Figure \ref{fig:different-sacles-tern}.} \textcolor{orange}{so we can visualize the trajectory of the epidemic in three-dimensional space (see Fig. \ref{fig:different-scales-tern} (left)).} For state space models like in our example, given the constraint that $S(t) + I(t)+R(t)$ is always equal to $N$ (the total population size), we can visual these point in a a two-dimesional \textit{ternary} plot, as seen in Figure \ref{fig:different-scales-tern} (right). \sout{In Fig. \ref{fig:different-scales-tern}
we plot the filamental trajectories of the two epidemics in a time-invariant view. We will explain how and why this works shortly. The important takeaway is that in this time-invariant view, i}I}t
is apparent that these epidemics are on ``the same path.'' In this case,
this indicates that two epidemics have the same value of \(R_0\).

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics[width=0.49\linewidth]{images/vis3d} \includegraphics[width=0.49\linewidth]{Figs/unnamed-chunk-3-2}

}

\caption{\label{fig:different-scales-tern}Left: trajectory of epidemic in three-dimensional space, plotting $(S(t), I(t), R(t))$.  Right: the gray-shaded region and epidemic trajectory shown from (left) now shown in two-dimensional space.  This is more commonly known as a ternary plot.}\label{fig:unnamed-chunk-3}
\end{figure}
\end{CodeChunk}

\textcolor{violet}{Underlying our time-invariant visualization that allowed for the comparison of $R_0$ in Fig. \ref{fig:different-scales-tern} is the treatment of the epidemic as a single filamental trajectory in the state space. \sout{The reason why we can visually compare $R_0$ in Fig. \ref{fig:different-scales-tern} is because of the time-invariant nature of the filamental trajectory associated with an epidemic.}}
A filamental trajectory can be mathematically viewed as a set of points
in space that have an ordering, and that all points on the line between
these ordered points are also contained in the geometric object. For a
SIR epidemic, we can represent the associated filamental trajectory
\(\psi\) as

\[
\psi = \left \{(S(t), I(t), R(t)): S, I, R \ge 0, S + I + R = N \right \}_{t\in[0,T]},
\] where a mapping \(\xi : s \to \mathbb{R}\) that is strictly
monotonically increasing would not change the definition of \(\psi\),
i.e.~\(\psi_\xi \equiv \psi\) where :

\[
\psi_\xi = \left \{(S(\xi(s)), I(\xi(s)), R(\xi(s))): S, I, R \ge 0, S + I + R = N \right \}_{s\in[0,T]}.
\] \textcolor{violet}{[Ben says: removed this paragraph now]} Since the
number \textcolor{orange}{of individuals} in each state is non-negative
and the sum over the three states for a given time point
\sout{sums to}\textcolor{orange}{is} \(N\), then
\textcolor{violet}{all points in} \(\psi\) will lay in a
\textbackslash ben\{two-dimensional triangular plane in
three-dimensional space.
\sout{We can then which can be visualize the \textcolor{orange}{full filamental trajectory} in a two-dimensional ternary plot}
As a result, we can visualize
\textcolor{orange}{the full filamental trajectory} of an epidemic in a
single
\textcolor{violet}{ternary plot \sout{2d plot and ultimately $R_0$}}.
\textcolor{orange}{[Shannon says: Show pic here?]}{]}\footnote{Which?  \textcolor{orange}{3d space one? But I'm leaning against it now.   3d never looks good in a paper.}}

\sout{\textcolor{violet}{[This section could be a bit less wordy. But generally good.]} We visualize the two epidemics in a global, ternary view in Figure \ref{fig:different-scales-tern}.  Without getting into too much detail of the intricacies of this plot, we immediately see the points of the two filaments $\psi$ seem to form the same trajectory.  Now, it is much clearer that \textcolor{violet}{\sout{Model} epidemic} 2 is following the same trajectory as \textcolor{violet}{\sout{Model} epidemic} 1 but is not as far along in the infection process. }

\sout{\textcolor{violet}{As suggested a few paragraphs back, t}}The
filamental trajectories in Fig. \ref{fig:different-scales-tern}
\sout{seem to overlap}, and we may suspect that something is
fundamentally linking these two different epidemics together.
Mathematically, we can show this fundamental link \sout{turns} is
\(R_0\). Let our two epidemics be presented as
\(\{(S_1(t), I_1(t), R_1(t))\}_{t\geq0}\),
\(\{(S_2(s), I_2(s), R_2(s))\}_{s \geq 0}\) respectively. As with the
example, assume both models have the same initial values
\((S(0), I(0), R(0))\), and let
\(R_0 =\frac{\beta_1}{\gamma_1} = \frac{\beta_2}{\gamma_2}\) where
\(\beta_i\) and \(\gamma_i\) are the average infection rate and recovery
rate, respectively, for SIR model \(i=1, 2\). And define \(a>0\) to be
the relative scalar such that \(\beta_2 = a \beta_1\) if and only if
\(\gamma_2 = a \gamma_1\).

\begin{theorem}\label{thm:sir-scale}
Let there be two SIR models as described above.  Then for all $t > 0$ there exists an $s>0$ such that $(S_1(t), I_1(t), R_1(t)) = (S_2(s), I_2(s), R_2(s))$.  Moreover, $s = \frac{1}{a}t$.
\end{theorem}

The proof of Theorem \ref{thm:sir-scale} relies on a fairly recent
result from \cite{Harko2014} and is shown in detail in Proof
\ref{proof:thm}. The consequence of Theorem \ref{thm:sir-scale} is that
for two SIR models that have the same initial percent of individuals in
each state and \(R_0\) then for every point on the epidemic path of the
first SIR model \textcolor{violet}{\sout{is also} can be mapped to} a
point on the epidemic path of the second SIR model.
\textcolor{orange}{In other words, the two epidemics form the same filamental trajectory. For SIR models with similar initial state percentages, time-invariant analysis allows practitioners to compare values of $R_0$ at a glance.}

\subsection[Beyond R0 and SIR]{Time-invariant analysis beyond \(R_0\)
and \sout{Kermack’s and McKendrick} SIR
Models\footnote{Probably will need to change this title...}}\label{sec:beyond-r0-sir}

Through the \(R_0\) example, we see that treating epidemics like
filamental trajectories embedded in a lower dimensional space allows us
to \sout{better}\textcolor{orange}{more fully} compare the overall
structure of the epidemic and see how the population is directly
impacted. Time-invariant tools \sout{that} can be useful even when the
underlying generative model for the epidemic is unknown or have more
than three epidemic states.

\textcolor{violet}{New paragraph} Viewing epidemics as filamental
trajectories provides \sout{a lot }new ways to compare and examine
epidemics in a time-invariant manner. \textbackslash ben\{For
\textcolor{orange}{completed?}epidemics that have ended, one way to
examine their filamental trajectories is to
\sout{redefine}\textcolor{orange}{represent} the
\textcolor{orange}{filamental} trajectory as a
\textcolor{orange}{finite sequence of equally spaced points.}
\sout{finite sequence of points on the filamental trajectory that are equally spaced (i.e. equa-distance between pairs of ordered points).}
\sout{For epidemics that have "played" themselves out, one way to represent their filamental trajectories to avoid \textcolor{orange}{confusion stemming from}\sout{impacts of} temporal structure is to define them as a sequence of points their trajectory with equi-distance between each point \textcolor{orange}{[Shannon says: are we missing a few words in this definition?]}}.
This representation induces a natural distance between this type of
representation between epidemics, specifically: \[
d_\text{equi-distance}(\psi_1, \psi_2)  = \int_{s \in [0,1]} (\psi_1'(s) - \psi_2'(s))^2 ds
\] where \(\psi_i'(s)\) the point along \(\psi_1\) that is \(s\)
fraction of \(|\psi_1|\) distance away from the start of
\(\psi_i\).\footnote{\textcolor{orange}{I think you're trying to say something about a distance based on the equally space points.  Some clarifying questions:  is $\psi^\prime$ the derivative?  Does proportion make more sense than fraction? or simply $\frac{|\psi|}{s}$? It's only naturally time-invariant if we have a well defined ending point, right?}}
This distance is naturally time-invariant, and can be plugged into
multiple distance-based assessment tools to examine the overall
``extremeness'' of points, including pseudo-density estimators and
depth/local depth functions
\citep[for examples see][]{Ciollaro2016, Geenens2017}. These extremeness
estimators can be \sout{very} useful when comparing
\sout{between a setof simulation} \textcolor{orange}{a set of simulated}
epidemics and the true
epidemic\textcolor{orange}{. Moreover these extremness estimators,}
\sout{and does} \textcolor{orange}{} not constrain the number of states
of the
model\sout{, though we recommend projecting the points into the unit simplex}\footnote{\textcolor{violet}{I'm not sure we've talked about this before... \textcolor{orange}{I don't think we have but am wondering if we're getting in the weeds}}}\sout{(by making all values the proportion of the population in the given state)}.

\textcolor{violet}{New paragraph:}
\textcolor{violet}{\sout{If the set of epidemics that one is examining have only gone through a single cycle of the outbreak} If \sout{one} \textcolor{orange}{a practicioner} is interested in understanding \textcolor{orange}{an} epidemic\sout{s} through a single \sout{cycle}\textcolor{orange}{realization} of \sout{their}\textcolor{orange}{its} outbreak}
(before the population of individuals have become susceptible again),
then additional time-invariant tools, including prediction regions can
be leveraged \textcolor{orange}{awk sentence}\footnote{\{\textcolor{orange}{What are we predicting if the epidemic is done? Update 4/6: I'm satisfied.}\}}.
In these settings, \textcolor{orange}{\pkg{EpiCompare} goes}
\sout{we go }a step further and treats epidemics more like geometric
filaments
\textcolor{orange}{(i.e. filamental trajectories without an ordering of points)}
than filamental trajectories. In \pkg{EpiCompare}, we create prediction
regions that contain a the top (\(1-\alpha\)) proportion of simulated
curves by defining geometric regions defined by the union of small
\textcolor{orange}{geometric?} filaments around the subset of
simulations (\sout{subset} \textcolor{orange}{grouped} by measures like
the above pseudo-density estimates or depth estimates). These regions
\sout{look at}\textcolor{orange}{show} where in the state-space we
expect the epidemic to traverse\sout{, and }.
\textcolor{orange}{Additionally, }we can compare prediction regions
defined by different models using
\textcolor{violet}{many set difference distances \sout{the Haussdorf \textcolor{orange}{why Haussdorf specifically?} distance}}
as well as examining how well the truth epidemic matches the simulations
by examining if the epidemic's trajectory lies within the prediction
region. All these \textcolor{orange}{mentioned?} geometric structures
and distance notations apply to epidemics with any number of states, and
at the end of Section \ref{sec:overview} we \sout{also} highlight how
these prediction regions can aid in visual comparisons for epidemics
with 3 states (like the SIR models).

\section[Package overview]{Overview of
\pkg{EpiCompare}}\label{sec:overview}

\afterpage{\clearpage}
\begin{sidewaysfigure}[!ht]
    \centering
    \includegraphics[width = 1\textwidth]{images/pipeline2_1.pdf}
    \caption{How \pkg{EpiCompare} supplements and aids in the epidemiological data analysis pipeline.}
    \label{fig:pipeline2}
\end{sidewaysfigure}

In this section, we present the tools implemented in \pkg{EpiCompare}
and explain how they aid in the data analysis pipeline. In Fig.
\ref{fig:pipeline2}, we illustrate how our package's functions fit into
the data analysis pipeline introduced in Fig. \ref{fig:pipeline}. All
front-facing functions are aimed to be as user-friendly as possible. We
also focus on providing the user ``tidyverse'' style functions, that
encourage piping objects from one function to the next and follow clear
``verb'' naming schemes \citep{Wickham2019}. Although users can
incorporate \pkg{EpiCompare} into any step in the data analysis
pipeline, there are two primary points of entry. The first point of
entry is the very beginning with pre-processing and visualizing raw
data, and the second point of entry is after modeling and simulation.
Figure \ref{fig:pipeline2} captures these different paths, and we
highlight\footnote{\textcolor{violet}{[Ben says: we need to make sure we actually do highlight ]}}
both approaches and how to leverage \pkg{EpiCompare} in the subsections
below.

\textbf{Data Pre-processing}

The first step of most data analysis is ``cleaning'' the data
\textcolor{orange}{to a format that is friendly for both computers and programmers}\footnote{\textcolor{violet}{[Ben says: I'm unsure this is needed and naturally adds more text. Willing to accept...]}}
so it can be explored.
\sout{There are multiple ways to collect epidemiological data.}
\textcolor{violet}{In epidemology, there are multiple different formats the data can arrive in.}
\footnote{\textcolor{orange}{[Shannon suggests: "Epidemiological data are collected in many different formats."]} \textcolor{violet}{[Ben says: I see the point you're trying to make and made a new edit - but it's less connected to the previous sentence since "format" is the end.]}}
Sometimes individual records are collected, with times of different
states of the epidemic (infection, recovery, etc.) as well as individual
information like network structure, location, and sub-population
information. Other data collections focus on aggregate counts of
individuals in each epidemic state. In fact, many times only the number
of new infections at each time step (e.g.~weekly case counts) is
observed. Compartment totals (amounts of individuals in each state) are
then imputed from those case
counts\sout{ along with}\textcolor{orange}{and}\footnote{\textcolor{violet}{[Ben says: Declined. I'm unclear why we should. Is the sentence unclear? Maybe it's the way it's a compound sentence? - If so a rewrite would be better?]}}
other information about the disease and the population of interest. In
\pkg{EpiCompare}, we focus on understanding the overall impact of an
outbreak at the aggregate/population level, which allows for streamlined
examination of overall trends of an epidemic.

To help the practitioner examine epidemics from an aggregate/ population
lens, we provide a function called \texttt{agents\_to\_aggregate()}.
This function transforms data about individual/agents' initial entry
into each state (e.g.~start of infection, start of recovery, etc.) to an
aggregate view of how many individuals were in a state at a given time.
There are often situations where
\textcolor{violet}{\sout{grouping} aggregating}\footnote{\textcolor{violet}{[Ben says: reason - connect to languare from sentence before.]}}
agents into subpopulations
(e.g.~sub\sout{populations}\textcolor{orange}{ groups}\footnote{\textcolor{violet}{[Ben says: declined. I'm unclear why we'd do this.]}}
defined by age or sex) can highlight different aggregate level trends.
For example, research by \citet{rvachev1985,anderson1992,worby2015}
develop state-based models that account for differing disease dynamics
in different subpopulations. In \pkg{EpiCompare}, we facilitate
subpopulation analysis by combining the function
\pkg{dplyr}::\texttt{group\_by()} and \texttt{agent\_to\_aggregate()} to
provide aggregation
\sout{at a}\textcolor{orange}{by}\footnote{\textcolor{violet}{[Ben says: declined. unclear about this change. Also would it not be "by groups"? Orignial seems to flow more.]}}
group level.

The \texttt{agents\_to\_aggregate()} function is flexible and can deal
with a wide range of information about each individual.
\textcolor{violet}{\sout{It can,} In fact, this function can} account
for infinitely many states. This functionality allows the practitioner
to aggregate information relative to common states
(e.g.~``Susceptible'', ``Infectious'', and ``Recovered'') as well as
more complex states (e.g.~``Exposed'', ``iMmune'', ``Hospitalized'').
Additionally, \texttt{agents\_to\_aggregate()} permits indicators for
death/exit and birth/entry dates. Overall, this function is a powerful
tool for pre-processing data,
\textcolor{orange}{and it lowers the barrier for entry into data analysis for less experienced practicioners.}\footnote{\textcolor{orange}{[Shannon says: I added this in theme with 'highlighting a point of entry']} \textcolor{violet}{[Ben says: I'm unclear about this comment but I'm find with the addition...]}}

\textbf{EDA}

\textcolor{violet}{\sout{With raw data, "Getting to know" our} In the early stages of a project, getting to know the}\footnote{\textcolor{orange}{[Shannon suggested: "Getting to know" our]} \textcolor{violet}{[Ben says: even though I was the original one with the quotes - one should always stray away from it. I kept the implicit connection to the section we're in given I'm not sure people are reading the titles super well.]}}
data currently means figuring out useful combinations of visualizations
\textcolor{violet}{\sout{,}and} numerical summaries and
\sout{\sout{subsets}\textcolor{orange}{groupings}}\textcolor{violet}{exploring different groupings of the data}\footnote{\textcolor{violet}{[Ben says: the action "figuring out" and "groupings" was a bit unclear what actually was going to happen - so I changed it.]}}.
\textcolor{violet}{\sout{An expert coder has many ways to successfully explore the data in an aggregate lens using }}\texttt{agents\_to\_aggregate()}\textcolor{violet}{\sout{. For less experienced coders, \pkg{EpiCompare} also includes tools to rapidly explore data that has three epidemic states.} An expert coder can start from }
\texttt{agents\_to\_aggregate()}
\textcolor{violet}{to successfully accomplish EDA in many ways, but \pkg{EpiCompare} also includes tools that allow a novice coder to rapidly explore data, as long as there three unique epidemiological states (like the SIR model).}\footnote{\textcolor{violet}{[Ben says: I'm reverting to the old text - it's much clearer. - happy to have a a discussion on it. The crossed out text didn't capture what an expert coder was really going to do and how that differed.]}}
Building on \textcolor{violet}{\sout{the tools in}} \pkg{ggplot2} and
\pkg{ggtern} \textcolor{violet}{packages}, \pkg{EpiCompare}`s
\texttt{geom\_aggregate()} provides a rapid way to explore different
subpopulations' experiences of an epidemic
\citep{Wickham2016, Hamilton2018}.
\textcolor{violet}{The function }\texttt{geom\_aggregate()}
\textcolor{violet}{provides a visualization tool to holistically examine aggregate level information across different subpopulations by visualizing each subpopulation's epidemic trajectory in the three-dimensional state space.}
\footnote{\textcolor{violet}{[Original text: "By combiing the ideas behind agent2aggregate for three-state models to examine subpopulation trajcetories in 2d simplex space."]} \textcolor{orange}{[Shannons says: I got rid of the end of this sentence because I think it was putting multiple ideas in one sentence.]}\textcolor{violet}{[Ben says: I'm unclear why this warrants just deleting it. I've provided a newly written sentence...]}}
Visualization tools for three-state models were developed because (1)
SIR models are some of the most common and basic epidemic state-based
models and (2) our three-dimensional simplex representation of these
epidemics emphasizes a ``time-invarance'' representation of the data
(for a refresher see Section \ref{sec:time-invariant}).

\textbf{Model Fitting and Simulations}

\textcolor{violet}{[Ben says: think about this section and if it highlights that we can bring in outside models...]}

After getting a good sense of what a past or current epidemic looks like
with EDA, the next step is often model fitting and/or simulations.
\sout{In this step and the next step (post-processing), we discuss how to \textcolor{orange}{easily include external models and simulations originating outside into the \pkg{EpiCompare} data analysis pipeline.}}\textcolor{violet}{In this step and the next step (post-proessing), we highlight how practitioners can pair models and simulations of epidemics from outside of \pkg{EpiCompare} with analysis and simulation tools in \pkg{EpiCompare}.}\footnote{\textcolor{violet}{[Ben says: the original rewrite didn't seem to clear - specifically in how the practitioner would be using outside models in this step.]}}
While \pkg{EpiCompare} does not focus on fitting model(s) to data, we do
provide some flexible functions for simulation of basic discrete-time
epidemic-state models.
\textcolor{violet}{These functions simulation individual-level information based on practitioner estimated transition rates between states and}
can be combined with \texttt{agents\_to\_aggregate()} to view these
simulations through an aggregate lens. The function
\texttt{simulate\_SIR\_agents()} simulates a basic SIR epidemic with
user inputs for the number of simulations, the initial number in each
state, the infection and recovery parameters \((\beta, \gamma\)), and
the total number of discrete time steps. This function allows for easy
access to SIR model analysis and comparison. Beyond SIR models, the
function \texttt{simulate\_agents()} takes as input a user-specified
state-transition matrix and other epidemic parameters to allow the user
to create simulations for an outbreak with \textit{any} number of states
and any number of transitions among them. This flexibility in states can
be used to also reflect group-based dynamics. In turn, this allows for
users to explore the space of models in an intuitive way without getting
bogged down by too much mathematical detail. For consistency, we have
made output from \texttt{simulate\_agents()} and
\texttt{simulate\_SIR\_agents()} compatible with
\texttt{agents\_to\_aggregate()} so aggregate information may easily be
accessed.

\textbf{Post-processing}

\textcolor{violet}{[Ben says: I think we should remind the reader that we care more about simulations, in order to compare fitted models between themselves and the true epidemics.]}

\textcolor{violet}{[Ben says: this replaces the first paragraph below]}

If the practitioner wishes to compare models-to-observations or even
models-to-models, they needs to post-process their models and
simulations to disseminate the results in an easily digestable format.
In general, post-processing of modeling and simulation consists of
making summary statistics, plots, and tables.
\textcolor{violet}{Model \sout{The}} summaries can be very complex, and
as a result, a number of epidemic modeling \textbackslash proglang\{R\}
packages return \textcolor{violet}{\sout{a}} special class
\textcolor{violet}{objects}. The special
class\textcolor{violet}{\sout{es} objects} often contain a plethora of
information from residuals, model diagnostics, input parameters, and
more. While incredibly useful, these special classes can be difficult
for novice coders to \textcolor{violet}{\sout{navigate} handle}.

To this end, \pkg{EpiCompare} provides a series of
\texttt{fortify}-style methods, called \texttt{fortify\_aggregate()}
which transform output from infectious disease modeling and simulation
packages like \pkg{pomp} and \pkg{EpiModel} into tidy-styled data frames
which contain information about the total number of individuals in each
state at a given time, for a given simulation. These fortify functions
have output that is consistent with that of
\texttt{agents\_to\_aggregate()}.

\textcolor{orange}{[Shannon says: Here Ben talks about filament compression and tidy\_dist mats, etc.]}
\textcolor{violet}{[Ben says: not currently thinking about including tidy\_dist stuff - will need to think about this as it is including in the pipeline 2 image...]}
\footnote{\textcolor{violet}{[Ben says: are these comments still relevant?]}}

To utilize simulations of epidemics in later time-invariant analysis we
also provide a function to convert temporally defined epidemics to
filamental representations. Specifically, we provide the function
\texttt{filament\_compression()} to convert simulation(s) to filaments
as expressed by presenting the epidemic as a ordering of some common
fixed number of points so that they are equally spaced along the
original path in the proportional state space.

\textcolor{orange}{[Shannon says: Shannon's attempt at above paragraph is below..  I tried to more smoothly connect to previous paragraph and then kinda copped out at trying to reword the filament description by just referring back to the previous section.]}

\pkg{EpiCompare} also provides a tool to convert time-dependent epidemic
simulations into their time-invariant filamental representations.
Specifically, \texttt{filament\_compression()} converts simulations to
filaments (see Section \ref{sec:time-invariant})
\textcolor{violet}{\sout{so practitioners can view model simulations and results through a time-invariant lens.}which allows practitioner to later apply time-invariant tools to these compressed epidemiological objects.}\footnote{\textcolor{violet}{[Ben says: I worry about this statement - it is very broad / seems to claim a lot but not very clearly...]}}
These tools were developed to provide another natural entry point into
the \pkg{EpiCompare} data analysis pipeline for situations where
modeling and analysis is already completed and practitioners are looking
for simple and transparent tools to help understand and disseminate
results.

\textbf{Comparisons and Assessment}

\textcolor{violet}{[Ben says: I didn't look at this section - it looks like your comments are more "let's" talk about this and not suggestions / feedback.]}

\textcolor{orange}{Finally, \pkg{EpiCompare} can be used the last step of the data analysis pipeline with its comparison and assessment utility functions.}
As introduced in Section \ref{sec:beyond-r0-sir} there's a
\sout{lot of}\textcolor{orange}{much} potential for time-invariant tools
to help compare and assess epidemics and models/simulations. In
\pkg{EpiCompare} we provide a set of comparison and assessment tools for
models that extend beyond the standard performance metrics (e.g.~mean
squared error or AIC) and focus on assessing the structural information
the models capture. This approaches work well on models where
\sout{online} one ``cycle'' of the epidemic has occurred (no recovered
individuals have been susceptible
again)\footnote{\textcolor{violet}{[Ben says: Shannon - do you think this is clear / a desirable way to define this - we define it slightly differently in section 2.2.]}\textcolor{orange}{Shannon says I don't think I understand what you are trying to say there.  We should chat about it.}}.
Epidemics are complex objects, and we provide tools to create prediction
regions
\textcolor{orange}{with differing desired characteristics}\footnote{\textcolor{orange}{Shannon says: I want to connect epidemics being complex to the fact that prediction is hard and not all prediction regions tell you the same thing}}
from simulated epidemics. For three-state epidemic models, we provide
the \texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion
{[}\textcolor{orange}{good place for Ben paper cite?}{]} of the
simulations (where the simulations treated as filaments). In
\pkg{EpiCompare} we also provide these prediction regions for epidemic
models with with more than three states. The functions
\texttt{create\_convex\_hull\_structure()} and
\texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any state-based model. For
both of these geometric objects, we provide functions to check if a path
is contained (\texttt{contained()}) and the ability to assess the
Hausdorff distance between prediction regions based on simulations from
different model (\texttt{hassdorf\_dist()}).
\footnote{\textcolor{violet}{[Ben says: should we discuss the details of geom\_prediction\_bands more? ]}\textcolor{orange}{[Shannon says: I don't think so, best left to the vignette and data example]}}

We also provide functions to calculate the ``extremeness'' of a true
epidemic {[}\textcolor{orange}{trajectory?}{]} compared to simulated
epidemics via the equi-distance filamental trajectory representation as
mentioned in Section \ref{sec:beyond-r0-sir}. Specifically, functions
like \texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.
\footnote{\textcolor{violet}{[Ben says: should we define these more in this paragraph? Introducted in 2.2. as well and not defined.]}\textcolor{orange}{[Shannon says: I'd throw a citation or two but again limit the detail.]}}

\textcolor{violet}{[Ben says: Replaced with paragraphs above.]}
Comparison and assessment of model fit or comparisons of one model to
another model can be performed in a variety of ways including mean
square error, AIC, plots, and more. Perhaps the most useful tool
\pkg{EpiCompare} has to offer to the expert, for comparison and
assessment of models, is in its post-processing tools which create a
standard output. It is then a matter of writing a script or function
made for that standard output to assess the results from multiple models
in the way the user
desires\footnote{\textcolor{violet}{[Ben says: Shannon, doesn't this below in the previous section?, I might drop this paragraph except for the introduction sentence - is it really true that the standardization is most valuable?]}\textcolor{orange}{I agree with you now that this section is nice and filled in.  Let's drop this paragraph.}}.

\textcolor{violet}{[Ben says: Replaced with paragraphs above.]} However,
for those who like more concrete tools, \pkg{EpiCompare} offers
functions to compare prediction regions to one another including
\texttt{geom\_prediction\_band()} (which plots the region), and
\texttt{create\_\{convex\_hull,delta\_ball\}\_structure()}(which returns
the \texttt{R} output for the given structure), and \texttt{contained()}
(which allows the user to determine if one set of points is contained in
a prediction band). Additionally, we offer ways to determine if model
outputs are compatible with one another, that is how extreme one output
is to another. \textcolor{orange}{Ben says something about distance}

\section[Tour]{A tour of \pkg{EpiCompare}}\label{sec:tour}

In this section, we highlight many of the tools available in
\pkg{EpiCompare}. As previously discussed, these tools include data
cleaning; visualization; modeling and simulation; post-processing; and
comparison and model assessment, in accordance with the data analysis
pipeline (Fig. \ref{fig:pipeline}). We show a full data analysis from
beginning to end that can be accomplished in a streamlined and
standardized manner via \pkg{EpiCompare}.

\subsection{Data and exploratory analysis}

We analyze an outbreak of measles in the town of Hagelloch, Germany from
1861-1862, a data set organized by \cite{pfeilsticker1863}. The data was
later made visible by \cite{oesterle1992} and made available in an
\proglang{R} by \cite{surveillance2017}. The Hagelloch data includes a
rich set of features including household members, school level,
household locations, date of first symptoms (prodromes), date of measles
rash, and even the alleged infector. A subset of the data is shown in
Table \ref{tab:hags-people}. Because of these rich features, this data
set has been an ideal testing ground methodology in infectious disease
epidemiology and is used in work by
\cite{Neal2004,britton2011,groendyke2012,becker2016}.

\begin{CodeChunk}
\begin{table}[!h]

\caption{\label{tab:hags-people}Subset of Hagelloch infection data.  Features include the person ID, household ID (HH ID), age, sex, class level (Pre-K/1st/2nd), date of first symptoms, date of the appearance of the measles rash, and the alleged infector ID of the individual.}
\centering
\begin{tabular}[t]{rrlrllllr}
\toprule
ID & HH ID & Name & Age & Sex & Class & Symp. Start & Rash Date & Infector ID\\
\midrule
1 & 61 & Mueller & 7 & female & 1st class & 1861-11-21 & 1861-11-25 & 45\\
2 & 61 & Mueller & 6 & female & 1st class & 1861-11-23 & 1861-11-27 & 45\\
3 & 61 & Mueller & 4 & female & preschool & 1861-11-28 & 1861-12-02 & 172\\
4 & 62 & Seibold & 13 & male & 2nd class & 1861-11-27 & 1861-11-28 & 180\\
5 & 63 & Motzer & 8 & female & 1st class & 1861-11-22 & 1861-11-27 & 45\\
45 & 51 & Goehring & 7 & male & 1st class & 1861-11-11 & 1861-11-13 & 184\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

With \pkg{EpiCompare}, we can easily obtain the empirical cumulative
incidence function with respect to the measles rash appearance (variable
\code{ERU}) with the following tidy-style function,
\code{agents_to_aggregate()}. The function \code{agents_to_aggregate()}
is a key component of \pkg{EpiCompare}, allowing the user to easily
switch from an individual-level (i.e.~an agent) view of a disease to an
aggregate level. For example, the below code shows how we can convert
the agent data to a cumulative incidence of the measles rash, in order
to see how the disease spread through the population over time. We can
then compare the cumulative incidence of the rash to the cumulative
incidence of the prodromes, i.e.~the initial symptoms. We do this with
the below code, and a part of the cumulative incidence data output is
shown in Table \ref{tab:cif-rash}. The argument
\code{integer_time_expansion} indicates whether we should include all
time points in the recorded range of the data or only when there is a
change in the incidence.

\begin{CodeChunk}
\begin{CodeInput}
R> cif_rash  <- hagelloch_raw %>%
+   mutate(time_of_rash = as.numeric(ERU - min(PRO, na.rm = TRUE))) %>%
+   agents_to_aggregate(states = time_of_rash,
+                       integer_time_expansion = FALSE) %>%
+   mutate(type = "Rash")
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{table}[!h]

\caption{\label{tab:cif-rash}Turning the individual-level information from the Hagelloch data to an aggregate view of the cumulative incidence of the measles rash in the population over time.}
\centering
\begin{tabular}[t]{rrr}
\toprule
Time & \# Susceptible & \# Total rash appearances\\
\midrule
0 & 188 & 0\\
4 & 187 & 1\\
7 & 186 & 2\\
9 & 185 & 3\\
12 & 183 & 5\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

One possible question of interest is the duration between initial onset
of prodromes and the appearance of the measles rash. Since
\code{agent_to_aggregate()} outputs a tidy-style data frame, it is a
simple task to plot the two sets of incidence curves on the same graph
(Fig. \ref{fig:cifs}).

\begin{CodeChunk}
\begin{CodeInput}
R> cif_prodromes <- hagelloch_raw %>%
+   mutate(time_of_PRO = as.numeric(PRO - min(PRO, na.rm = TRUE))) %>%
+   agents_to_aggregate(states = time_of_PRO,
+                       integer_time_expansion = FALSE) %>%
+   mutate(type = "Pro")
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> plot_df <- bind_rows(cif_rash, cif_prodromes)
R> 
R> ggplot(data = plot_df,
+        aes(x = t, y = X1, col = type)) + 
+   geom_step() + 
+   labs(title = "Cumulative incidence of measles appearance",
+        x = "Time (days relative to first prodrome appearance)",
+        y = "Cumulative incidence of event") + 
+   coord_cartesian(xlim = c(0, 55)) +
+   scale_color_manual(values = c("blue", "red"))
\end{CodeInput}
\begin{figure}[H]

{\centering \includegraphics{Figs/unnamed-chunk-7-1}

}

\caption{\label{fig:cifs}Empirical cumulative incidence functions of prodrome (symptom) onset and measles rash appearance.  We see that there is approximately a a constant lag between the two curves.}\label{fig:unnamed-chunk-7}
\end{figure}
\end{CodeChunk}

The real power of \code{agents_to_aggregate()} lies in its ability to
aggregate over any number of pre-specified states. For example, the
Hagelloch data sets contains two columns, \code{tI} and \code{tR}, the
time of infection and recovery, respectively of each individual. We can
then plot the SIR values through a time-invariant lens using
\pkg{ggplot2} and \pkg{ggtern} functions (as shown in Fig.
\ref{fig:hag-tern-raw}) or with our custom \code{geom},
\code{geom_aggregate}, which takes the raw agent data as input.

\begin{CodeChunk}
\begin{CodeInput}
R> hagelloch_sir <- hagelloch_raw %>%
+   agents_to_aggregate(states = c(tI, tR),
+                       min_max_time = c(0, 55)) %>%
+   rename(time = t, S = X0, I = X1, R = X2)
R> 
R> 
R> ggplot(hagelloch_sir, aes(x = S, y = I, z = R))+
+   coord_tern() +
+   geom_path() +
+   labs(x = "S", y = "I", z = "R",
+        title = "Time invariant view of Hagelloch measles outbreak") + 
+   theme_sir(base_size = 24)
\end{CodeInput}
\begin{figure}[H]

{\centering \includegraphics{Figs/unnamed-chunk-9-1}

}

\caption{\label{fig:hag-tern-raw}Time invariant view of the Hagelloch epidemic where we view the individuals in Susceptible, Infectious, or Recovered states.  We see there are two peaks of infection (the vertical axis).}\label{fig:unnamed-chunk-9}
\end{figure}
\end{CodeChunk}

Moreover, we can look at the outbreaks of the disease by group within
\code{agent_to_aggregate()} or \code{geom_aggregate()}. This allows us
to examine differences among the different groups of individuals. For
example, we show the time invariant outbreak by class level in Figure
\ref{fig:tern-class-data}. Immediately, we see that time invariant
infection curve is different for the pre-school class compared to the
1st class. In the 1st class, we see about 95\% of the class become
infected and less than 10\% of them having recovered, which may be
indicative of a super-spreading event. This suspicion is further
confirmed in that 26 of the 30 1st class students have been reportedly
infected by the same individual.

\begin{CodeChunk}
\begin{CodeInput}
R> hagelloch_raw %>%
+   ggplot(aes(y = tI, z = tR, color = CL)) +
+   geom_aggregate(size = 2) + coord_tern() +
+   labs(x = "S", y = "I", z = "R",
+        color = "Class") +
+   scale_color_brewer(palette = "Dark2") +
+   facet_wrap(~CL)
\end{CodeInput}
\begin{figure}[H]

{\centering \includegraphics{Figs/unnamed-chunk-10-1}

}

\caption{\label{fig:tern-class-data}Time invariant outbreak curves for the three class groups.  The pre-school class has a distinct peak of infection whereas the peak infection point for the other two classes are less well defined.}\label{fig:unnamed-chunk-10}
\end{figure}
\end{CodeChunk}

Along with multiple epidemic states, the function
\code{agents_to_aggregate} can also be extended to populations with
vital dynamics (e.g.~birth and death) and examples of this are shown in
the package vignette. In summary, \code{agents_to_aggregate()} is a
multi-purpose workhorse that may be leveraged to convert individual
level records into aggregate information that may be more useful for
some forms of epidemic modeling such as compartment modeling.

Up to this point, we have used \pkg{EpiCompare} in the context of
observed data. We also want to compare statistical models, and
\pkg{EpiCompare} aids in that process via a simple yet flexible
individual-level simulator, conversion tools for popular epidemic model
packages, and model assessments. We demonstrate an example here.

We first try to model the Hagelloch data with a stochastic SIR model,
which we refer to as the `simple SIR.' In our vignette, we show how to
fit this simple SIR model via maximum likelihood and simulate from the
model with those best fit parameters. Our function
\code{simulate_agents()} generates individual level data according to
discrete time multinomial draws, which depend on the number of
individuals in each state at the previous time step and a matrix of
transition probabilities. For example, the below code generates 100
simulations of an outbreak of a disease with one initial infector in a
population of \(n= 188\) individuals.

\begin{CodeChunk}
\begin{CodeInput}
R> trans_mat <- matrix(c("X0 * (1 - X1 * par1 / N)", "X0 * X1  * par1 / N", "0",
+                   "0", "X1 * (1 - par2)", "par2 * X1",
+                   "0", "0", "X2"), byrow = TRUE, nrow = 3)
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> set.seed(2020)
R> 
R> best_params <- c("beta" = .36, "gamma" = .13)
R> ## This is the SIR representation
R> 
R> rownames(trans_mat) <- c("S", "I", "R")
R> init_vals <- c(187, 1, 0)
R> par_vals <- c(par1 = best_params[1], par2 = best_params[2])
R> max_T <- 55
R> n_sims <- 100
R> 
R> agents <- simulate_agents(trans_mat,
+                        init_vals,
+                        par_vals,
+                        max_T,
+                        n_sims,
+                        verbose = FALSE)
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> agg_model <- agents %>% group_by(sim) %>%
+   agents_to_aggregate(states = c(I, R)) %>%
+   mutate(Type = "Simple SIR")
\end{CodeInput}
\end{CodeChunk}

The result of our simulation is the object \code{agents} which is a
18800 \(\times\) 5 data frame, which details the time of entry into the
\(S\), \(I\), and \(R\) states for a given simulation. Before we examine
the results of this simple SIR model, we will also examine another, more
sophisticated SIR model, this time from the package \pkg{EpiModel}.
Briefly, this model first fits a contact network to the set of
indivduals, where the class of the student is a covariate. The model
then simulates a SIR-epidemic on that network.

\begin{CodeChunk}
\begin{CodeInput}
R> library(EpiModel)
R> ## WARNING:  Will take a minute or two
R> 
R> set.seed(42)
R> nw <- network.initialize(n = 188, directed = FALSE)
R> nw <- set.vertex.attribute(nw, "group", rep(0:2, each = 90, 30, 68))
R> formation <- ~edges + nodematch("group") + concurrent
R> target.stats <- c(200, 300, 200)
R> coef.diss <- dissolution_coefs(dissolution = ~offset(edges),  duration = 5)
R> est1 <- netest(nw, formation, target.stats, coef.diss, edapprox = TRUE)
R> 
R> param <- param.net(inf.prob = 0.1, act.rate = 5, rec.rate = 0.1)
R> status.vector <- c(rep(0, 90), rep(0, 30), rep(0, 67), 1)
R> status.vector <- ifelse(status.vector == 1, "i", "s")
R> init <- init.net(status.vector = status.vector)
R> control <- control.net(type = "SIR", nsteps = 55,
+                        nsims = 100, epi.by = "group")
R> epimodel_sir <- netsim(est1, param, init, control)
\end{CodeInput}
\end{CodeChunk}

The output of this model is \code{epimodel_sir}, an object of class
\code{netsim}, which contains a plethora of modeling information. We
provide the function \code{fortify_aggregate()}, which can take objects
from specialized classes of modeling output and transform it into a
tidy-style data frame.

\begin{CodeChunk}
\begin{CodeInput}
R> fortified_net <- fortify_aggregate(epimodel_sir, 
+                                    states = c("s.num", "i.num", "r.num")) %>%
+   mutate(Type = "EpiModel SIR",
+          sim = as.numeric(gsub("sim", "", sim)))
\end{CodeInput}
\end{CodeChunk}

We can then analyze the results of the two models side by side as
time-invariant epidemic curves. The results are shown in Figure
\ref{fig:hag-simple-sir}, where a 90\% prediction band is estimated from
the delta ball method for each of the two models. For the Simple SIR
model, we see that the data generally covers the data fairly well but
clearly misses the second peak of infection. We also see that the
prediction band is very large, covering up a large area of the ternary
plot. On the other hand, for the \pkg{EpiModel} model, we see that the
prediction band covers the data quite well and takes up less area.

\begin{CodeChunk}
\begin{CodeInput}
R> both_models <- bind_rows(agg_model, fortified_net)
R> 
R> 
R> g <- ggplot() + geom_prediction_band(data = both_models %>% filter(t != 0),
+          aes(x = X0, y = X1, z = X2,
+               sim_group = sim, fill = Type),
+          alpha = .5,
+          conf_level = .90) 
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> g +   geom_path(data = both_models %>% filter(t !=0),
+             aes(x = X0, y = X1, z = X2, group = paste(Type, sim)),
+             alpha = .3, col = "gray40") + 
+     coord_tern() + theme_sir(base_size = 24) +
+   geom_point(data = hagelloch_sir,
+              aes(x = S, y = I, z =R), col = "black") +
+   labs(title = "Simple SIR model",
+        subtitle = "90% Prediction band and original data",
+        x = "S", y = "I", z = "R") +
+   scale_fill_manual(values = c("#006677", "#AA6600")) + 
+   facet_wrap(~Type) +
+   theme(legend.position = "bottom")
\end{CodeInput}
\begin{figure}[H]

{\centering \includegraphics{Figs/unnamed-chunk-17-1}

}

\caption{\label{fig:hag-simple-sir}  Original Hagelloch SIR data (black) along with 90\% prediction band and actual simulation paths from the Simple SIR and the EpiModel SIR models.}\label{fig:unnamed-chunk-17}
\end{figure}
\end{CodeChunk}

However, both models are not a good fit to the filamental path as
opposed to the individual points in \((S, I, R)\)-space. This can be
captured with the set of simulations both models predict (gray lines),
which all generally have a single defined peak of infection whereas the
data certainly looks like it has two distinct peaks, likely caused by
our assumed super-spreader event. This observation is backed up by the
below analysis that demonstrates that the estimated pseudo-density of
the observed epidemic (relative to the simulations from either model) is
much less likely then \textbf{any} of the simulations (reported in Table
\ref{tab:hags-extreme}). In conclusion, \pkg{EpiCompare} makes it clear
that, at a glance, 1) the \pkg{EpiModel} network model is a better fit
than the Simple SIR model, and 2) the fit is only good at the geometric
filamental level as opposed to the epidemic trajectory filamental level.

\begin{CodeChunk}
\begin{CodeInput}
R> #-- after cleaning up and combining --
R> all_together_df <- rbind(simple_sir,
+                          hagelloch_sir2)
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{table}[!h]

\caption{\label{tab:cif-all-together-df}Top and bottom 2 rows of \tt{all\_together\_df}\text{, combining both simulated epidemics and the true epidemic.}}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
Type & sim & t & S & I & R\\
\midrule
Simple SIR & 1 & 0 & 188 & 0 & 0\\
Simple SIR & 1 & 1 & 187 & 1 & 0\\
true observation & 0 & 54 & 1 & 0 & 187\\
true observation & 0 & 55 & 1 & 0 & 187\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> compression_df <- all_together_df %>% group_by(Type, sim) %>% 
+   filament_compression(data_columns = c("S","I","R"), 
+                        number_points = 20)
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> tdmat <- compression_df %>% 
+   dist_matrix_innersq_direction(
+     position = c(1:length(compression_df))[
+       names(compression_df) %in% c("S","I", "R")],
+     tdm_out = T)
R> 
R> simple_sir_true_obs_info <- tdmat %>% 
+   compare_new_to_rest_via_distance(
+     new_name_id = data.frame(Type = "true observation", sim = 0),
+     distance_func = distance_psuedo_density_function, 
+     sigma = "20%") 
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{table}[!h]

\caption{\label{tab:hags-extreme}The extremeness of the true simulations based on comparing psuedo-density estimates between true vs simulated curves}
\centering
\begin{tabular}[t]{l>{\raggedleft\arraybackslash}p{6cm}>{\raggedleft\arraybackslash}p{6cm}}
\toprule
Type & simulations-based estimated psuedo-density & proportion of simulations with lower estimated psuedo-density\\
\midrule
Simple SIR & 0.0036733 & 0\\
EpiModel SIR & 0.0028813 & 0\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

Overall, \pkg{EpiCompare} aids in the data analysis pipeline for both
novice and expert practitioners and coders alike. These tools encourage
model and simulation exploration of many of the existing and
well-supported packages that already exist, and side-by-side comparison
thereof. Finally, we hope that practicioners will consider using
time-invariant analysis when trying to assess and compare epidemics and
epidemic models.

\hypertarget{a.-appendix}{%
\section*{A. Appendix}\label{a.-appendix}}
\addcontentsline{toc}{section}{A. Appendix}

\hypertarget{a.1-proof-of-theorem}{%
\subsection*{\texorpdfstring{A.1 Proof of Theorem
\ref{thm:sir-scale}}{A.1 Proof of Theorem }}\label{a.1-proof-of-theorem}}
\addcontentsline{toc}{subsection}{A.1 Proof of Theorem
\ref{thm:sir-scale}}

\begin{proof}\label{proof:thm}
\cite{Harko2014} provide an analytical solution for the Kermack and McKendrick equations (Eq. \eqref{eq:sir-ode}) by reparameterizing the ODEs so that $\mathcal{S}(u) = S(t)$, $\mathcal{I}(u) = S(t)$, and $\mathcal{R}(u) = R(t)$ for $0< u_T < 1$ with
\begin{align}\label{eq:harko-odes}
\mathcal{S}(u) &= S(0)u\\
\mathcal{I}(u) &= N - R(0) + NR_0^{-1}\log u - S(0)u \nonumber\\
\mathcal{R}(u) &= R(0) - NR_0^{-1} \log u, \nonumber
\end{align}
and $u$ and t are related by the following integral,
\begin{align*}
    t &= \int_{u}^1 \frac{N}{\beta \tau (N - R(0) + R_{0}^{-1} \log \tau - S(0)\tau)}d\tau \\
    &= \int_{u}^1 \frac{1}{\beta f(S(0), R(0), N, R_0, \tau)} d \tau\\
    &= \int_{u}^1 \frac{1}{\beta f(\tau)} d\tau,
\end{align*}
where we have made the denominator of the integral a function of $N$, the initial values, $R_0$, and $\tau$, which we further condense to $f(\tau)$ for brevity.
Then for a given $t$ we want to find $s$ such that $(S_1(t), I_1(t), R_1(t)) = (S_2(s), I_2(s), R_2(s))$.  Or equivalently, for a fixed $u$ want to find $v$ such that  $\mathcal{S}_1(u) = \mathcal{S}_2(v)$ and then the corresponding $t$ and $s$ are given by
\begin{align*}
    t & = \int_{u}^1 \frac{1}{\beta_1 f(\tau)} d\tau \\
    s & = \int_{v}^1 \frac{1}{\beta_2 f(\tau)} d\tau.
\end{align*}
Note that since the equations in Eq. \eqref{eq:harko-odes} are functions of the initial values and $R_0$, then $u = v$. We then can find a relation for $s$,
    \begin{align*}
    s & = \int_{u}^1 \frac{1}{\beta_2 f(\tau)} d\tau  \\
    & = \int_{u}^1 \frac{1}{a\beta_1 f(\tau)} d\tau \\ 
    &= \frac{1}{a}\int_{u}^1 \frac{1}{\beta_1 f(\tau)} d\tau \\
    &= \frac{1}{a}t.
\end{align*}
\end{proof}

\bibliography{EpiCompare.bib}

\end{document}